---
title: "Single Polyp Stats"
author: "Shayle Matsuda"
date: "10/28/2021"
output:
  pdf_document: default
  word_document: default
  html_document: 
    self_contained: no
editor_options:
  chunk_output_type: console
---
PERMANOVA for single polyp metabolomics

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
```

# libraries
```{r,echo=FALSE}  
# knitr::opts_chunk$set(warning=FALSE, message=FALSE)
library(ape)
library(MuMIn)
library(plyr) 
library(dplyr)
library(ggplot2)
library(ggpubr)
library(lmerTest)
library(car)
library(emmeans)
library(gridExtra)
library(multcomp)
library(reshape)
library(tidyverse)
library(factoextra)
library(reshape2)
library(vegan) 
library(pairwiseAdonis)
library("scales")
packageVersion("scales")
library(RColorBrewer)
library(colorRamps)
library(devtools)
library(phyloseq)
library(readr)
library(microbiome)
library(vegan)
```

Load in data:
```{r}
mData<-read.csv("single polyp master with PCoA coordinates.csv") #metabolomics data and xy coords from Ty
mData$ATTRIBUTE_Sample_Name<-as.factor(as.character(mData$ATTRIBUTE_Sample_Name))
mData$ATTRIBUTE_Colony_number<-as.factor(as.character(mData$ATTRIBUTE_Colony_number))
mData<-subset(mData, ATTRIBUTE_Colony_number!="976B")#remove dup

```



# PERMANOVA
```{r}
#PERMANOVA

#sep dfs
mData.sd<-mData[,1:11] #metadata
mData.meta<-mData[,12:566] #metabolomics data
 #mData.PCoA<-mData[,567:568] #ty's x and y, but can't use these for STATS
 #plot(mData.PCoA) #plot Ty's x and y, to confirm your distance matrix cooresponds right

#write.csv(mData.sd,"mData.sd3.csv")
####################### Between colonies and Between Sample points


# Bray-Curtis distance 
mData.dist <- vegdist(mData[,12:566],  method = "bray")
mData.dist.mat<-as.matrix(mData.dist)
 #write.csv(mData.dist.mat, "mData.dist.bc.csv")

PCOA <- pcoa(mData.dist) #pcoa
barplot(PCOA$values$Relative_eig[1:10]) # plot the eigenvalues 
biplot.pcoa(PCOA) #plot
PCOAaxes <- PCOA$vectors[,c(1,2)] #for visualization only

mData.dist2<-as.matrix(mData.dist)

 mData.dist.sd<- inner_join(rownames_to_column(mData.sd), rownames_to_column(data.frame(PCOAaxes)), type = "right", by = "rowname") #merge PCOAaxes and metadata

 
 BetweenColonies_plot<-ggplot(mData.dist.sd, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(colour = Symbiont.Type), size = 4) +
  #scale_color_manual(values=Sp.colors)+
  #geom_text(label = rownames(scores), nudge_x = 0.05, nudge_y = 0.05,check_overlap = T) +
 # stat_ellipse(aes(x = Axis.1, y = Axis.2, colour = ATTRIBUTE_Colony_number), linetype = 2) +
  ggtitle("PCoA Between Colonies") +
  theme_classic();BetweenColonies_plot
 
 metaMDS()
 
 
 
############ plot Between colonies ####

BetweenColonies_plot<-ggplot(mData.dist.sd, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(colour = ATTRIBUTE_Colony_number), size = 4) +
  #scale_color_manual(values=Sp.colors)+
  #geom_text(label = rownames(scores), nudge_x = 0.05, nudge_y = 0.05,check_overlap = T) +
 # stat_ellipse(aes(x = Axis.1, y = Axis.2, colour = ATTRIBUTE_Colony_number), linetype = 2) +
  ggtitle("PCoA Between Colonies") +
  theme_classic();BetweenColonies_plot
       
#PERMANOVA between COLONIES
set.seed(30)
adonis2(mData.dist~ATTRIBUTE_Colony_number, data=mData.sd)

#                          Df SumsOfSqs MeanSqs F.Model      R2 Pr(>F)   
# ATTRIBUTE_Colony_number  18    18.732 1.04064  32.641 0.64527  0.001 ***
# Residuals               323    10.297 0.03188         0.35473           
# Total                   341    29.029                 1.00000     



############# Between Sampling areas (polyps) ########
set.seed(30)
adonis2(mData.dist~Area, data=mData.sd)
#            Df SumsOfSqs  MeanSqs F.Model      R2 Pr(>F)  
# Area        5    0.7291 0.145811  1.7312 0.02511  0.017 *
# Residuals 336   28.3000 0.084226         0.97489         
# Total     341   29.0290                  1.00000    

mData.sd$PolypZone2<-mData.sd$Area #duplicate col
mData.sd$PolypZone2[mData.sd$PolypZone2 == "Touching 1"] <- "Base"
mData.sd$PolypZone2[mData.sd$PolypZone2 == "Touching 2"] <- "Base"
mData.sd$PolypZone2[mData.sd$PolypZone2 == "Midle 1"] <- "Mid"
mData.sd$PolypZone2[mData.sd$PolypZone2 == "Midle 2"] <- "Mid"

adonis2(mData.dist~PolypZone2, data=mData.sd)
pairwise.adonis2(mData.dist, mData.sd$PolypZone2)

mData.sd$Area[mData.sd$Area == "Base"] <- "1"
mData.sd$Area[mData.sd$Area == "Touching 1"] <- "2"
mData.sd$Area[mData.sd$Area == "Touching 2"] <- "3"
mData.sd$Area[mData.sd$Area == "Midle 1"] <- "4"
mData.sd$Area[mData.sd$Area == "Midle 2"] <- "5"
mData.sd$Area[mData.sd$Area == "Tip"] <- "6"

pairwise.adonis2(mData.dist, mData.sd$Area)



adonis2(mData.dist~Area*ATTRIBUTE_Colony_number, data=mData.sd)
#                               Df SumsOfSqs MeanSqs F.Model      R2 Pr(>F)    
# Area                           5    0.7291 0.14581   4.864 0.02511  0.001 ***
# ATTRIBUTE_Colony_number       18   18.7315 1.04064  34.712 0.64527  0.001 ***
# Area:ATTRIBUTE_Colony_number  90    2.7332 0.03037   1.013 0.09415  0.445    
# Residuals                    228    6.8353 0.02998         0.23546           
# Total                        341   29.0290                 1.00000    

adonis2(mData.dist~ATTRIBUTE_Colony_number*ATTRIBUTE_Branch_Letter*Area, data=mData.sd)
# 
#                                                       Df SumsOfSqs MeanSqs F.Model      R2 Pr(>F)
# Area                                                   5    0.7291       0       0 0.02511      1
# ATTRIBUTE_Colony_number                               18   18.7315       1       0 0.64527      1
# ATTRIBUTE_Branch_Letter                                2    0.1206       0       0 0.00415      1
# Area:ATTRIBUTE_Colony_number                          90    2.7332       0       0 0.09415      1
# Area:ATTRIBUTE_Branch_Letter                          10    0.2231       0       0 0.00768      1
# ATTRIBUTE_Colony_number:ATTRIBUTE_Branch_Letter       36    1.7009       0       0 0.05859      1
# Area:ATTRIBUTE_Colony_number:ATTRIBUTE_Branch_Letter 180    4.7907       0       0 0.16503      1
# Residuals                                              0    0.0000    -Inf         0.00000       
# Total                                                341   29.0290                 1.00000       


adonis2(mData.dist~ATTRIBUTE_Branch_Letter*Area, data=mData.sd)




#Plot: No patterns
BetweenSampArea_plot<-ggplot(mData.dist.sd, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(colour = Area), size = 4) +
  #scale_color_manual(values=Sp.colors)+
  #geom_text(label = rownames(scores), nudge_x = 0.05, nudge_y = 0.05,check_overlap = T) +
 # stat_ellipse(aes(x = Axis.1, y = Axis.2, colour = ATTRIBUTE_Colony_number), linetype = 2) +
  ggtitle("PCoA Between Sampling Area") +
  theme_classic();BetweenSampArea_plot

```
Between polyps 1-3 only
```{r}
# ########################### Between polyps 1-3 only
#subset data
mData.Polyps123<-subset(mData, Area=="Base"|Area=="Touching 1"|Area=="Touching 2")
mData.Polyps123.sd<-mData.Polyps123[,1:11]
# Bray-Curtis distance 
mData.Polyps123.dist <- vegdist(mData.Polyps123[,12:566],  method = "bray")

PCOA <- pcoa(mData.Polyps123.dist) #pcoa
barplot(PCOA$values$Relative_eig[1:10]) # plot the eigenvalues 
biplot.pcoa(PCOA) #plot
PCOAaxes <- PCOA$vectors[,c(1,2)] #for visualization only

 mData.Polyps123.dist.sd<- inner_join(rownames_to_column(mData.Polyps123.sd), rownames_to_column(data.frame(PCOAaxes)), type = "right", by = "rowname") #merge PCOAaxes and metadata


set.seed(30)
adonis2(mData.Polyps123.dist~Area, data=mData.Polyps123.sd)
#            Df SumsOfSqs  MeanSqs F.Model      R2 Pr(>F)
# Area        2    0.0603 0.030125 0.33161 0.00393  0.992
# Residuals 168   15.2620 0.090846         0.99607       
# Total     170   15.3223                  1.00000 


set.seed(30)
adonis2(mData.Polyps123.dist~Area*ATTRIBUTE_Colony_number, data=mData.Polyps123.sd)
#                               Df SumsOfSqs MeanSqs F.Model      R2 Pr(>F)    
# Area                           2    0.0603 0.03013  0.9495 0.00393  0.460    
# ATTRIBUTE_Colony_number       18   10.7537 0.59743 18.8308 0.70183  0.001 ***
# Area:ATTRIBUTE_Colony_number  36    0.8916 0.02477  0.7806 0.05819  0.983    
# Residuals                    114    3.6168 0.03173         0.23605           
# Total                        170   15.3223                 1.00000         


#Plot by Colony number (see that still cluster by colony)
BetweenPoly123_plot<-ggplot(mData.Polyps123.dist.sd, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(colour = ATTRIBUTE_Colony_number), size = 4) +
  #scale_color_manual(values=Sp.colors)+
  #geom_text(label = rownames(scores), nudge_x = 0.05, nudge_y = 0.05,check_overlap = T) +
 # stat_ellipse(aes(x = Axis.1, y = Axis.2, colour = ATTRIBUTE_Colony_number), linetype = 2) +
  ggtitle("PCoA Between Polyps 1-3") +
  theme_classic();BetweenPoly123_plot

#Plot by Polyp number (see that no cluster)
BetweenPoly123_plot2<-ggplot(mData.Polyps123.dist.sd, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(colour = Area), size = 4) +
  #scale_color_manual(values=Sp.colors)+
  #geom_text(label = rownames(scores), nudge_x = 0.05, nudge_y = 0.05,check_overlap = T) +
 # stat_ellipse(aes(x = Axis.1, y = Axis.2, colour = ATTRIBUTE_Colony_number), linetype = 2) +
  ggtitle("PCoA Between Polyps 1-3 by polyp#") +
  theme_classic();BetweenPoly123_plot2


```
Between Branches by individual colony
```{r}
#Between Branches by colony ###########################



####### colony 945  ####### 
mData.Branch945<-subset(mData, ATTRIBUTE_Colony_number=="945") 
#subset data
mData.Branch945.sd<-mData.Branch945[,1:11]
# Bray-Curtis distance 
mData.Branch945.dist <- vegdist(mData.Branch945[,12:566],  method = "bray")

PCOA <- pcoa(mData.Branch945.dist) #pcoa
barplot(PCOA$values$Relative_eig[1:10]) # plot the eigenvalues 
biplot.pcoa(PCOA) #plot
PCOAaxes <- PCOA$vectors[,c(1,2)] #for visualization only

 mData.Branch945.dist.sd<- inner_join(rownames_to_column(mData.Branch945.sd), rownames_to_column(data.frame(PCOAaxes)), type = "right", by = "rowname") #merge PCOAaxes and metadata


set.seed(30)
adonis2(mData.Branch945.dist~Area, data=mData.Branch945.sd)
#            Df SumsOfSqs  MeanSqs F.Model      R2 Pr(>F)
# Area       5   0.16096 0.032193 0.96577 0.28694  0.499
# Residuals 12   0.40001 0.033334         0.71306       
# Total     17   0.56097                  1.00000    

#Plot by Colony number (see that still cluster by colony)
Branch945_plot<-ggplot(mData.Branch945.dist.sd, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(colour = ATTRIBUTE_Branch_Letter), size = 2) +
  #scale_color_manual(values=Sp.colors)+
  #geom_text(label = rownames(scores), nudge_x = 0.05, nudge_y = 0.05,check_overlap = T) +
 # stat_ellipse(aes(x = Axis.1, y = Axis.2, colour = ATTRIBUTE_Colony_number), linetype = 2) +
  ggtitle("PCoA Between Branches 945") +
  theme_classic();Branch945_plot

# test diff between polyps within branch
#subset data
mData.Branch945.sd<-mData.Branch945[,1:11]
# Bray-Curtis distance 
mData.Branch945.dist <- vegdist(mData.Branch945[,12:566],  method = "bray")

PCOA <- pcoa(mData.Branch945.dist) #pcoa
barplot(PCOA$values$Relative_eig[1:10]) # plot the eigenvalues 
biplot.pcoa(PCOA) #plot
PCOAaxes <- PCOA$vectors[,c(1,2)] #for visualization only

 mData.Branch945.dist.sd<- inner_join(rownames_to_column(mData.Branch945.sd), rownames_to_column(data.frame(PCOAaxes)), type = "right", by = "rowname") #merge PCOAaxes and metadata


set.seed(30)
adonis2(mData.Branch945.dist~Area, data=mData.Branch945.sd)
#            Df SumsOfSqs  MeanSqs F.Model      R2 Pr(>F)
# Area       5   0.16096 0.032193 0.96577 0.28694  0.499
# Residuals 12   0.40001 0.033334         0.71306       
# Total     17   0.56097                  1.00000    

#Plot by Colony number (see that still cluster by colony)
Branch945_plot<-ggplot(mData.Branch945.dist.sd, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(colour = ATTRIBUTE_Branch_Letter), size = 2) +
  #scale_color_manual(values=Sp.colors)+
  #geom_text(label = rownames(scores), nudge_x = 0.05, nudge_y = 0.05,check_overlap = T) +
 # stat_ellipse(aes(x = Axis.1, y = Axis.2, colour = ATTRIBUTE_Colony_number), linetype = 2) +
  ggtitle("PCoA Between Branches 945") +
  theme_classic();Branch945_plot



####### colony 959  ####### 
mData.Branch959<-subset(mData, ATTRIBUTE_Colony_number=="959") 
#subset data
mData.Branch959.sd<-mData.Branch959[,1:11]
# Bray-Curtis distance 
mData.Branch959.dist <- vegdist(mData.Branch959[,12:566],  method = "bray")

PCOA <- pcoa(mData.Branch959.dist) #pcoa
barplot(PCOA$values$Relative_eig[1:10]) # plot the eigenvalues 
biplot.pcoa(PCOA) #plot
PCOAaxes <- PCOA$vectors[,c(1,2)] #for visualization only

 mData.Branch959.dist.sd<- inner_join(rownames_to_column(mData.Branch959.sd), rownames_to_column(data.frame(PCOAaxes)), type = "right", by = "rowname") #merge PCOAaxes and metadata


set.seed(30)
adonis2(mData.Branch959.dist~Area, data=mData.Branch959.sd)
# #            Df SumsOfSqs  MeanSqs F.Model      R2 Pr(>F)
# Area       5  0.084667 0.016933  1.1214 0.31845  0.303
# Residuals 12  0.181202 0.015100         0.68155       
# Total     17  0.265869                  1.00000       

#Plot by Colony number (see that still cluster by colony)
Branch959_plot<-ggplot(mData.Branch959.dist.sd, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(colour = ATTRIBUTE_Branch_Letter), size = 2) +
  #scale_color_manual(values=Sp.colors)+
  #geom_text(label = rownames(scores), nudge_x = 0.05, nudge_y = 0.05,check_overlap = T) +
 # stat_ellipse(aes(x = Axis.1, y = Axis.2, colour = ATTRIBUTE_Colony_number), linetype = 2) +
  ggtitle("PCoA Between Branches 959") +
  theme_classic();Branch959_plot

####### colony 960  ####### 
mData.Branch960<-subset(mData, ATTRIBUTE_Colony_number=="960") 
#subset data
mData.Branch960.sd<-mData.Branch960[,1:11]
# Bray-Curtis distance 
mData.Branch960.dist <- vegdist(mData.Branch960[,12:566],  method = "bray")

PCOA <- pcoa(mData.Branch960.dist) #pcoa
barplot(PCOA$values$Relative_eig[1:10]) # plot the eigenvalues 
biplot.pcoa(PCOA) #plot
PCOAaxes <- PCOA$vectors[,c(1,2)] #for visualization only

 mData.Branch960.dist.sd<- inner_join(rownames_to_column(mData.Branch960.sd), rownames_to_column(data.frame(PCOAaxes)), type = "right", by = "rowname") #merge PCOAaxes and metadata


set.seed(30)
adonis2(mData.Branch960.dist~ATTRIBUTE_Branch_Letter, data=mData.Branch960.sd)
# #            Df SumsOfSqs  MeanSqs F.Model      R2 Pr(>F)
# ATTRIBUTE_Branch_Letter  2   0.05525 0.027625  1.4696 0.16384  0.164
# Residuals               15   0.28196 0.018798         0.83616       
# Total                   17   0.33721                  1.00000          

#Plot by Colony number (see that still cluster by colony)
Branch960_plot<-ggplot(mData.Branch960.dist.sd, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(colour = ATTRIBUTE_Branch_Letter), size = 2) +
  #scale_color_manual(values=Sp.colors)+
  #geom_text(label = rownames(scores), nudge_x = 0.05, nudge_y = 0.05,check_overlap = T) +
 # stat_ellipse(aes(x = Axis.1, y = Axis.2, colour = ATTRIBUTE_Colony_number), linetype = 2) +
  ggtitle("PCoA Between Branches 960") +
  theme_classic();Branch960_plot

####### colony 961  Sig####### 
mData.Branch961<-subset(mData, ATTRIBUTE_Colony_number=="961") 
#subset data
mData.Branch961.sd<-mData.Branch961[,1:11]
# Bray-Curtis distance 
mData.Branch961.dist <- vegdist(mData.Branch961[,12:566],  method = "bray")

PCOA <- pcoa(mData.Branch961.dist) #pcoa
barplot(PCOA$values$Relative_eig[1:10]) # plot the eigenvalues 
biplot.pcoa(PCOA) #plot
PCOAaxes <- PCOA$vectors[,c(1,2)] #for visualization only

 mData.Branch961.dist.sd<- inner_join(rownames_to_column(mData.Branch961.sd), rownames_to_column(data.frame(PCOAaxes)), type = "right", by = "rowname") #merge PCOAaxes and metadata


set.seed(30)
adonis2(mData.Branch961.dist~ATTRIBUTE_Branch_Letter, data=mData.Branch961.sd)
#           Df SumsOfSqs  MeanSqs F.Model      R2 Pr(>F)
# ATTRIBUTE_Branch_Letter  2   0.12052 0.060262  2.7776 0.27026   0.03 *
# Residuals               15   0.32543 0.021695         0.72974         
# Total                   17   0.44595                  1.00000              

#Plot by Colony number (see that still cluster by colony)
Branch961_plot<-ggplot(mData.Branch961.dist.sd, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(colour = ATTRIBUTE_Branch_Letter), size=2) +
  #scale_color_manual(values=Sp.colors)+
  #geom_text(label = rownames(scores), nudge_x = 0.05, nudge_y = 0.05,check_overlap = T) +
 # stat_ellipse(aes(x = Axis.1, y = Axis.2, colour = ATTRIBUTE_Colony_number), linetype = 2) +
  ggtitle("PCoA Between Branches 961") +
  theme_classic();Branch961_plot

####### colony 962 Significant ####### 
mData.Branch962<-subset(mData, ATTRIBUTE_Colony_number=="962") 
#subset data
mData.Branch962.sd<-mData.Branch962[,1:11]
# Bray-Curtis distance 
mData.Branch962.dist <- vegdist(mData.Branch962[,12:566],  method = "bray")

PCOA <- pcoa(mData.Branch962.dist) #pcoa
barplot(PCOA$values$Relative_eig[1:10]) # plot the eigenvalues 
biplot.pcoa(PCOA) #plot
PCOAaxes <- PCOA$vectors[,c(1,2)] #for visualization only

 mData.Branch962.dist.sd<- inner_join(rownames_to_column(mData.Branch962.sd), rownames_to_column(data.frame(PCOAaxes)), type = "right", by = "rowname") #merge PCOAaxes and metadata


set.seed(30)
adonis2(mData.Branch962.dist~ATTRIBUTE_Branch_Letter, data=mData.Branch962.sd)
# ATTRIBUTE_Branch_Letter  2   0.23235 0.11618  2.3623 0.23953  0.016 *
# Residuals               15   0.73770 0.04918         0.76047         
# Total                   17   0.97005                 1.00000         

#Plot by Colony number (see that still cluster by colony)
Branch962_plot<-ggplot(mData.Branch962.dist.sd, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(colour = ATTRIBUTE_Branch_Letter), size=2) +
  #scale_color_manual(values=Sp.colors)+
  #geom_text(label = rownames(scores), nudge_x = 0.05, nudge_y = 0.05,check_overlap = T) +
 # stat_ellipse(aes(x = Axis.1, y = Axis.2, colour = ATTRIBUTE_Colony_number), linetype = 2) +
  ggtitle("PCoA Between Branches 962") +
  theme_classic();Branch962_plot

####### colony 963 Sig ####### 
mData.Branch963<-subset(mData, ATTRIBUTE_Colony_number=="963") 
#subset data
mData.Branch963.sd<-mData.Branch963[,1:11]
# Bray-Curtis distance 
mData.Branch963.dist <- vegdist(mData.Branch963[,12:566],  method = "bray")

PCOA <- pcoa(mData.Branch963.dist) #pcoa
barplot(PCOA$values$Relative_eig[1:10]) # plot the eigenvalues 
biplot.pcoa(PCOA) #plot
PCOAaxes <- PCOA$vectors[,c(1,2)] #for visualization only

 mData.Branch963.dist.sd<- inner_join(rownames_to_column(mData.Branch963.sd), rownames_to_column(data.frame(PCOAaxes)), type = "right", by = "rowname") #merge PCOAaxes and metadata


set.seed(30)
adonis2(mData.Branch963.dist~ATTRIBUTE_Branch_Letter, data=mData.Branch963.sd)
# #            Df SumsOfSqs  MeanSqs F.Model      R2 Pr(>F)
# ATTRIBUTE_Branch_Letter  2  0.072242 0.036121  4.0202 0.34897  0.004 **
# Residuals               15  0.134771 0.008985         0.65103          
# Total                   17  0.207013                  1.00000          

#Plot by Colony number (see that still cluster by colony)
Branch963_plot<-ggplot(mData.Branch963.dist.sd, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(colour = ATTRIBUTE_Branch_Letter), size=2) +
  #scale_color_manual(values=Sp.colors)+
  #geom_text(label = rownames(scores), nudge_x = 0.05, nudge_y = 0.05,check_overlap = T) +
 # stat_ellipse(aes(x = Axis.1, y = Axis.2, colour = ATTRIBUTE_Colony_number), linetype = 2) +
  ggtitle("PCoA Between Branches 963") +
  theme_classic();Branch963_plot

####### colony 964  ####### 
mData.Branch964<-subset(mData, ATTRIBUTE_Colony_number=="964") 
#subset data
mData.Branch964.sd<-mData.Branch964[,1:11]
# Bray-Curtis distance 
mData.Branch964.dist <- vegdist(mData.Branch964[,12:566],  method = "bray")

PCOA <- pcoa(mData.Branch964.dist) #pcoa
barplot(PCOA$values$Relative_eig[1:10]) # plot the eigenvalues 
biplot.pcoa(PCOA) #plot
PCOAaxes <- PCOA$vectors[,c(1,2)] #for visualization only

 mData.Branch964.dist.sd<- inner_join(rownames_to_column(mData.Branch964.sd), rownames_to_column(data.frame(PCOAaxes)), type = "right", by = "rowname") #merge PCOAaxes and metadata


set.seed(30)
adonis2(mData.Branch964.dist~ATTRIBUTE_Branch_Letter, data=mData.Branch964.sd)
# #            Df SumsOfSqs  MeanSqs F.Model      R2 Pr(>F)
# ATTRIBUTE_Branch_Letter  2   0.06878 0.034389 0.56486 0.07004  0.901
# Residuals               15   0.91322 0.060881         0.92996       
# Total                   17   0.98200                  1.00000       

#Plot by Colony number (see that still cluster by colony)
Branch964_plot<-ggplot(mData.Branch964.dist.sd, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(colour = ATTRIBUTE_Branch_Letter), size=2) +
  #scale_color_manual(values=Sp.colors)+
  #geom_text(label = rownames(scores), nudge_x = 0.05, nudge_y = 0.05,check_overlap = T) +
 # stat_ellipse(aes(x = Axis.1, y = Axis.2, colour = ATTRIBUTE_Colony_number), linetype = 2) +
  ggtitle("PCoA Between Branches 964") +
  theme_classic();Branch964_plot

####### colony 965  ####### 
mData.Branch965<-subset(mData, ATTRIBUTE_Colony_number=="965") 
#subset data
mData.Branch965.sd<-mData.Branch965[,1:11]
# Bray-Curtis distance 
mData.Branch965.dist <- vegdist(mData.Branch965[,12:566],  method = "bray")

PCOA <- pcoa(mData.Branch965.dist) #pcoa
barplot(PCOA$values$Relative_eig[1:10]) # plot the eigenvalues 
biplot.pcoa(PCOA) #plot
PCOAaxes <- PCOA$vectors[,c(1,2)] #for visualization only

 mData.Branch965.dist.sd<- inner_join(rownames_to_column(mData.Branch965.sd), rownames_to_column(data.frame(PCOAaxes)), type = "right", by = "rowname") #merge PCOAaxes and metadata


set.seed(30)
adonis2(mData.Branch965.dist~ATTRIBUTE_Branch_Letter, data=mData.Branch965.sd)
# # #            Df SumsOfSqs  MeanSqs F.Model      R2 Pr(>F)
# ATTRIBUTE_Branch_Letter  2  0.035052 0.017526  1.6975 0.18456  0.096 .
# Residuals               15  0.154870 0.010325         0.81544         
# Total                   17  0.189922                  1.00000         

#Plot by Colony number (see that still cluster by colony)
Branch965_plot<-ggplot(mData.Branch965.dist.sd, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(colour = ATTRIBUTE_Branch_Letter), size=2) +
  #scale_color_manual(values=Sp.colors)+
  #geom_text(label = rownames(scores), nudge_x = 0.05, nudge_y = 0.05,check_overlap = T) +
 # stat_ellipse(aes(x = Axis.1, y = Axis.2, colour = ATTRIBUTE_Colony_number), linetype = 2) +
  ggtitle("PCoA Between Branches 965") +
  theme_classic();Branch965_plot

####### colony 966  ####### 
mData.Branch966<-subset(mData, ATTRIBUTE_Colony_number=="966") 
#subset data
mData.Branch966.sd<-mData.Branch966[,1:11]
# Bray-Curtis distance 
mData.Branch966.dist <- vegdist(mData.Branch966[,12:566],  method = "bray")

PCOA <- pcoa(mData.Branch966.dist) #pcoa
barplot(PCOA$values$Relative_eig[1:10]) # plot the eigenvalues 
biplot.pcoa(PCOA) #plot
PCOAaxes <- PCOA$vectors[,c(1,2)] #for visualization only

 mData.Branch966.dist.sd<- inner_join(rownames_to_column(mData.Branch966.sd), rownames_to_column(data.frame(PCOAaxes)), type = "right", by = "rowname") #merge PCOAaxes and metadata


set.seed(30)
adonis2(mData.Branch966.dist~ATTRIBUTE_Branch_Letter, data=mData.Branch966.sd)
# #            Df SumsOfSqs  MeanSqs F.Model      R2 Pr(>F)
# ATTRIBUTE_Branch_Letter  2   0.09030 0.045150  1.6735 0.18243  0.063 .
# Residuals               15   0.40469 0.026979         0.81757         
# Total                   17   0.49499                  1.00000  

#Plot by Colony number (see that still cluster by colony)
Branch966_plot<-ggplot(mData.Branch966.dist.sd, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(colour = ATTRIBUTE_Branch_Letter), size=2) +
  #scale_color_manual(values=Sp.colors)+
  #geom_text(label = rownames(scores), nudge_x = 0.05, nudge_y = 0.05,check_overlap = T) +
 # stat_ellipse(aes(x = Axis.1, y = Axis.2, colour = ATTRIBUTE_Colony_number), linetype = 2) +
  ggtitle("PCoA Between Branches 966") +
  theme_classic();Branch966_plot



####### colony 967 Sig  ####### 
mData.Branch967<-subset(mData, ATTRIBUTE_Colony_number=="967") 
#subset data
mData.Branch967.sd<-mData.Branch967[,1:11]
# Bray-Curtis distance 
mData.Branch967.dist <- vegdist(mData.Branch967[,12:566],  method = "bray")

PCOA <- pcoa(mData.Branch967.dist) #pcoa
barplot(PCOA$values$Relative_eig[1:10]) # plot the eigenvalues 
biplot.pcoa(PCOA) #plot
PCOAaxes <- PCOA$vectors[,c(1,2)] #for visualization only

 mData.Branch967.dist.sd<- inner_join(rownames_to_column(mData.Branch967.sd), rownames_to_column(data.frame(PCOAaxes)), type = "right", by = "rowname") #merge PCOAaxes and metadata


set.seed(30)
adonis2(mData.Branch967.dist~ATTRIBUTE_Branch_Letter, data=mData.Branch967.sd)
# #            Df SumsOfSqs  MeanSqs F.Model      R2 Pr(>F)
# ATTRIBUTE_Branch_Letter  2  0.063923 0.031961  2.1131 0.21981  0.007 **
# Residuals               15  0.226880 0.015125         0.78019          
# Total                   17  0.290802                  1.00000 

#Plot by Colony number (see that still cluster by colony)
Branch967_plot<-ggplot(mData.Branch967.dist.sd, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(colour = ATTRIBUTE_Branch_Letter), size=2) +
  #scale_color_manual(values=Sp.colors)+
  #geom_text(label = rownames(scores), nudge_x = 0.05, nudge_y = 0.05,check_overlap = T) +
 # stat_ellipse(aes(x = Axis.1, y = Axis.2, colour = ATTRIBUTE_Colony_number), linetype = 2) +
  ggtitle("PCoA Between Branches 967") +
  theme_classic();Branch967_plot

####### colony 968  Sig #######  
mData.Branch968<-subset(mData, ATTRIBUTE_Colony_number=="968") 
#subset data
mData.Branch968.sd<-mData.Branch968[,1:11]
# Bray-Curtis distance 
mData.Branch968.dist <- vegdist(mData.Branch968[,12:566],  method = "bray")

PCOA <- pcoa(mData.Branch968.dist) #pcoa
barplot(PCOA$values$Relative_eig[1:10]) # plot the eigenvalues 
biplot.pcoa(PCOA) #plot
PCOAaxes <- PCOA$vectors[,c(1,2)] #for visualization only

 mData.Branch968.dist.sd<- inner_join(rownames_to_column(mData.Branch968.sd), rownames_to_column(data.frame(PCOAaxes)), type = "right", by = "rowname") #merge PCOAaxes and metadata


set.seed(30)
adonis2(mData.Branch968.dist~ATTRIBUTE_Branch_Letter, data=mData.Branch968.sd)
# # #            Df SumsOfSqs  MeanSqs F.Model      R2 Pr(>F)
# ATTRIBUTE_Branch_Letter  2   0.13431 0.067156  1.9566 0.2069   0.04 *
# Residuals               15   0.51484 0.034323         0.7931         
# Total                   17   0.64915                  1.0000    

#Plot by Colony number (see that still cluster by colony)
Branch968_plot<-ggplot(mData.Branch968.dist.sd, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(colour = ATTRIBUTE_Branch_Letter), size=2) +
  #scale_color_manual(values=Sp.colors)+
  #geom_text(label = rownames(scores), nudge_x = 0.05, nudge_y = 0.05,check_overlap = T) +
 # stat_ellipse(aes(x = Axis.1, y = Axis.2, colour = ATTRIBUTE_Colony_number), linetype = 2) +
  ggtitle("PCoA Between Branches 968") +
  theme_classic();Branch968_plot

####### colony 969  ####### 
mData.Branch969<-subset(mData, ATTRIBUTE_Colony_number=="969") 
#subset data
mData.Branch969.sd<-mData.Branch969[,1:11]
# Bray-Curtis distance 
mData.Branch969.dist <- vegdist(mData.Branch969[,12:566],  method = "bray")

PCOA <- pcoa(mData.Branch969.dist) #pcoa
barplot(PCOA$values$Relative_eig[1:10]) # plot the eigenvalues 
biplot.pcoa(PCOA) #plot
PCOAaxes <- PCOA$vectors[,c(1,2)] #for visualization only

 mData.Branch969.dist.sd<- inner_join(rownames_to_column(mData.Branch969.sd), rownames_to_column(data.frame(PCOAaxes)), type = "right", by = "rowname") #merge PCOAaxes and metadata


set.seed(30)
adonis2(mData.Branch969.dist~ATTRIBUTE_Branch_Letter, data=mData.Branch969.sd)
# # #            Df SumsOfSqs  MeanSqs F.Model      R2 Pr(>F)
# ATTRIBUTE_Branch_Letter  2   0.05642 0.028210  1.1902 0.13695  0.276
# Residuals               15   0.35554 0.023703         0.86305       
# Total                   17   0.41196                  1.00000       

#Plot by Colony number (see that still cluster by colony)
Branch969_plot<-ggplot(mData.Branch969.dist.sd, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(colour = ATTRIBUTE_Branch_Letter), size=2) +
  #scale_color_manual(values=Sp.colors)+
  #geom_text(label = rownames(scores), nudge_x = 0.05, nudge_y = 0.05,check_overlap = T) +
 # stat_ellipse(aes(x = Axis.1, y = Axis.2, colour = ATTRIBUTE_Colony_number), linetype = 2) +
  ggtitle("PCoA Between Branches 969") +
  theme_classic();Branch969_plot

####### colony 971  ####### 
mData.Branch971<-subset(mData, ATTRIBUTE_Colony_number=="971") 
#subset data
mData.Branch971.sd<-mData.Branch971[,1:11]
# Bray-Curtis distance 
mData.Branch971.dist <- vegdist(mData.Branch971[,12:566],  method = "bray")

PCOA <- pcoa(mData.Branch971.dist) #pcoa
barplot(PCOA$values$Relative_eig[1:10]) # plot the eigenvalues 
biplot.pcoa(PCOA) #plot
PCOAaxes <- PCOA$vectors[,c(1,2)] #for visualization only

 mData.Branch971.dist.sd<- inner_join(rownames_to_column(mData.Branch971.sd), rownames_to_column(data.frame(PCOAaxes)), type = "right", by = "rowname") #merge PCOAaxes and metadata


set.seed(30)
adonis2(mData.Branch971.dist~ATTRIBUTE_Branch_Letter, data=mData.Branch971.sd)
# #            Df SumsOfSqs  MeanSqs F.Model      R2 Pr(>F)
# ATTRIBUTE_Branch_Letter  2   0.07725 0.038623  1.1665 0.1346  0.339
# Residuals               15   0.49665 0.033110         0.8654       
# Total                   17   0.57390                  1.0000     

#Plot by Colony number (see that still cluster by colony)
Branch971_plot<-ggplot(mData.Branch971.dist.sd, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(colour = ATTRIBUTE_Branch_Letter), size=2) +
  #scale_color_manual(values=Sp.colors)+
  #geom_text(label = rownames(scores), nudge_x = 0.05, nudge_y = 0.05,check_overlap = T) +
 # stat_ellipse(aes(x = Axis.1, y = Axis.2, colour = ATTRIBUTE_Colony_number), linetype = 2) +
  ggtitle("PCoA Between Branches 971") +
  theme_classic();Branch971_plot

####### colony 974 Sig  ####### 
mData.Branch974<-subset(mData, ATTRIBUTE_Colony_number=="974") 
#subset data
mData.Branch974.sd<-mData.Branch974[,1:11]
# Bray-Curtis distance 
mData.Branch974.dist <- vegdist(mData.Branch974[,12:566],  method = "bray")

PCOA <- pcoa(mData.Branch974.dist) #pcoa
barplot(PCOA$values$Relative_eig[1:10]) # plot the eigenvalues 
biplot.pcoa(PCOA) #plot
PCOAaxes <- PCOA$vectors[,c(1,2)] #for visualization only

 mData.Branch974.dist.sd<- inner_join(rownames_to_column(mData.Branch974.sd), rownames_to_column(data.frame(PCOAaxes)), type = "right", by = "rowname") #merge PCOAaxes and metadata


set.seed(30)
adonis2(mData.Branch974.dist~ATTRIBUTE_Branch_Letter, data=mData.Branch974.sd)
# #            Df SumsOfSqs  MeanSqs F.Model      R2 Pr(>F)
# ATTRIBUTE_Branch_Letter  2  0.065853 0.032926  3.0197 0.28705  0.001 ***
# Residuals               15  0.163557 0.010904         0.71295           
# Total                   17  0.229409                  1.00000          

#Plot by Colony number (see that still cluster by colony)
Branch974_plot<-ggplot(mData.Branch974.dist.sd, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(colour = ATTRIBUTE_Branch_Letter), size=2) +
  #scale_color_manual(values=Sp.colors)+
  #geom_text(label = rownames(scores), nudge_x = 0.05, nudge_y = 0.05,check_overlap = T) +
 # stat_ellipse(aes(x = Axis.1, y = Axis.2, colour = ATTRIBUTE_Colony_number), linetype = 2) +
  ggtitle("PCoA Between Branches 974") +
  theme_classic();Branch974_plot

####### colony 976  ####### 
mData.Branch976<-subset(mData, ATTRIBUTE_Colony_number=="976") 
#subset data
mData.Branch976.sd<-mData.Branch976[,1:11]
# Bray-Curtis distance 
mData.Branch976.dist <- vegdist(mData.Branch976[,12:566],  method = "bray")

PCOA <- pcoa(mData.Branch976.dist) #pcoa
barplot(PCOA$values$Relative_eig[1:10]) # plot the eigenvalues 
biplot.pcoa(PCOA) #plot
PCOAaxes <- PCOA$vectors[,c(1,2)] #for visualization only

 mData.Branch976.dist.sd<- inner_join(rownames_to_column(mData.Branch976.sd), rownames_to_column(data.frame(PCOAaxes)), type = "right", by = "rowname") #merge PCOAaxes and metadata


set.seed(30)
adonis2(mData.Branch976.dist~ATTRIBUTE_Branch_Letter, data=mData.Branch976.sd)
# #            Df SumsOfSqs  MeanSqs F.Model      R2 Pr(>F)
# ATTRIBUTE_Branch_Letter  2   0.19834 0.099168  1.6206 0.17769  0.128
# Residuals               15   0.91786 0.061191         0.82231       
# Total                   17   1.11620                  1.00000       

#Plot by Colony number (see that still cluster by colony)
Branch976_plot<-ggplot(mData.Branch976.dist.sd, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(colour = ATTRIBUTE_Branch_Letter), size=2) +
  #scale_color_manual(values=Sp.colors)+
  #geom_text(label = rownames(scores), nudge_x = 0.05, nudge_y = 0.05,check_overlap = T) +
 # stat_ellipse(aes(x = Axis.1, y = Axis.2, colour = ATTRIBUTE_Colony_number), linetype = 2) +
  ggtitle("PCoA Between Branches 976") +
  theme_classic();Branch976_plot

####### colony 978  ####### 
mData.Branch978<-subset(mData, ATTRIBUTE_Colony_number=="978") 
#subset data
mData.Branch978.sd<-mData.Branch978[,1:11]
# Bray-Curtis distance 
mData.Branch978.dist <- vegdist(mData.Branch978[,12:566],  method = "bray")

PCOA <- pcoa(mData.Branch978.dist) #pcoa
barplot(PCOA$values$Relative_eig[1:10]) # plot the eigenvalues 
biplot.pcoa(PCOA) #plot
PCOAaxes <- PCOA$vectors[,c(1,2)] #for visualization only

 mData.Branch978.dist.sd<- inner_join(rownames_to_column(mData.Branch978.sd), rownames_to_column(data.frame(PCOAaxes)), type = "right", by = "rowname") #merge PCOAaxes and metadata


set.seed(30)
adonis2(mData.Branch978.dist~ATTRIBUTE_Branch_Letter, data=mData.Branch978.sd)
# #            Df SumsOfSqs  MeanSqs F.Model      R2 Pr(>F)
# ATTRIBUTE_Branch_Letter  2   0.05551 0.027757  1.2815 0.14593   0.27
# Residuals               15   0.32490 0.021660         0.85407       
# Total                   17   0.38042                  1.00000    

#Plot by Colony number (see that still cluster by colony)
Branch978_plot<-ggplot(mData.Branch978.dist.sd, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(colour = ATTRIBUTE_Branch_Letter), size=2) +
  #scale_color_manual(values=Sp.colors)+
  #geom_text(label = rownames(scores), nudge_x = 0.05, nudge_y = 0.05,check_overlap = T) +
 # stat_ellipse(aes(x = Axis.1, y = Axis.2, colour = ATTRIBUTE_Colony_number), linetype = 2) +
  ggtitle("PCoA Between Branches 978") +
  theme_classic();Branch978_plot

####### colony 983  ####### 
mData.Branch983<-subset(mData, ATTRIBUTE_Colony_number=="983") 
#subset data
mData.Branch983.sd<-mData.Branch983[,1:11]
# Bray-Curtis distance 
mData.Branch983.dist <- vegdist(mData.Branch983[,12:566],  method = "bray")

PCOA <- pcoa(mData.Branch983.dist) #pcoa
barplot(PCOA$values$Relative_eig[1:10]) # plot the eigenvalues 
biplot.pcoa(PCOA) #plot
PCOAaxes <- PCOA$vectors[,c(1,2)] #for visualization only

 mData.Branch983.dist.sd<- inner_join(rownames_to_column(mData.Branch983.sd), rownames_to_column(data.frame(PCOAaxes)), type = "right", by = "rowname") #merge PCOAaxes and metadata


set.seed(30)
adonis2(mData.Branch983.dist~ATTRIBUTE_Branch_Letter, data=mData.Branch983.sd)
# #            Df SumsOfSqs  MeanSqs F.Model      R2 Pr(>F)
# ATTRIBUTE_Branch_Letter  2   0.21713 0.108565  1.0892 0.12681  0.367
# Residuals               15   1.49516 0.099677         0.87319       
# Total                   17   1.71229                  1.00000

#Plot by Colony number (see that still cluster by colony)
Branch983_plot<-ggplot(mData.Branch983.dist.sd, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(colour = ATTRIBUTE_Branch_Letter), size=2) +
  #scale_color_manual(values=Sp.colors)+
  #geom_text(label = rownames(scores), nudge_x = 0.05, nudge_y = 0.05,check_overlap = T) +
 # stat_ellipse(aes(x = Axis.1, y = Axis.2, colour = ATTRIBUTE_Colony_number), linetype = 2) +
  ggtitle("PCoA Between Branches 983") +
  theme_classic();Branch983_plot

####### colony 984  ####### 
mData.Branch984<-subset(mData, ATTRIBUTE_Colony_number=="984") 
#subset data
mData.Branch984.sd<-mData.Branch984[,1:11]
# Bray-Curtis distance 
mData.Branch984.dist <- vegdist(mData.Branch984[,12:566],  method = "bray")

PCOA <- pcoa(mData.Branch984.dist) #pcoa
barplot(PCOA$values$Relative_eig[1:10]) # plot the eigenvalues 
biplot.pcoa(PCOA) #plot
PCOAaxes <- PCOA$vectors[,c(1,2)] #for visualization only

 mData.Branch984.dist.sd<- inner_join(rownames_to_column(mData.Branch984.sd), rownames_to_column(data.frame(PCOAaxes)), type = "right", by = "rowname") #merge PCOAaxes and metadata


set.seed(30)
adonis2(mData.Branch984.dist~ATTRIBUTE_Branch_Letter, data=mData.Branch984.sd)
# #            Df SumsOfSqs  MeanSqs F.Model      R2 Pr(>F)
# ATTRIBUTE_Branch_Letter  2  0.044803 0.022402  1.7273 0.18719  0.091 .
# Residuals               15  0.194538 0.012969         0.81281         
# Total                   17  0.239341                  1.00000     

#Plot by Colony number (see that still cluster by colony)
Branch984_plot<-ggplot(mData.Branch984.dist.sd, aes(x = Axis.1, y = Axis.2)) + 
  geom_point(aes(colour = ATTRIBUTE_Branch_Letter), size=2) +
  #scale_color_manual(values=Sp.colors)+
  #geom_text(label = rownames(scores), nudge_x = 0.05, nudge_y = 0.05,check_overlap = T) +
 # stat_ellipse(aes(x = Axis.1, y = Axis.2, colour = ATTRIBUTE_Colony_number), linetype = 2) +
  ggtitle("PCoA Between Branches 984") +
  theme_classic();Branch984_plot


```

Within branches
```{r}

####### colony 957  ####### #not working yet
#subset branch A
mData.Branch957.A<-subset(mData.Branch957, ATTRIBUTE_Branch_Letter=="A") 
#subset data
mData.Branch957.A.sd<-mData.Branch957.A[,1:11]
# Bray-Curtis distance 
mData.Branch957.A.dist <- vegdist(mData.Branch957.A[,12:566],  method = "bray")

PCOA <- pcoa(mData.Branch957.A.dist) #pcoa
barplot(PCOA$values$Relative_eig[1:10]) # plot the eigenvalues 
biplot.pcoa(PCOA) #plot
PCOAaxes <- PCOA$vectors[,c(1,2)] #for visualization only

 mData.Branch957.A.dist.sd<- inner_join(rownames_to_column(mData.Branch957.A.sd), rownames_to_column(data.frame(PCOAaxes)), type = "right", by = "rowname") #merge PCOAaxes and metadata


set.seed(30)
adonis2(mData.Branch957.A.dist~Area, data=mData.Branch957.A.sd)
#           Df SumsOfSqs MeanSqs F.Model R2 Pr(>F)
# Area       5  0.097114       0       0  1      1
# Residuals  0  0.000000    -Inf          0       
# Total      5  0.097114                  1       









```



Plot all Colonies in same plot
```{r}
###### All plots together
Branch945_plot<-Branch945_plot+ theme(legend.position = "none") #remove legend
Branch957_plot<-Branch957_plot+ theme(legend.position = "none") #remove legend
Branch959_plot<-Branch959_plot+ theme(legend.position = "none") #remove legend
Branch960_plot<-Branch960_plot+ theme(legend.position = "none") #remove legend
Branch961_plot<-Branch961_plot+ theme(legend.position = "none") #remove legend
Branch962_plot<-Branch962_plot+ theme(legend.position = "none") #remove legend
Branch963_plot<-Branch963_plot+ theme(legend.position = "none") #remove legend
Branch964_plot<-Branch964_plot+ theme(legend.position = "none") #remove legend
Branch965_plot<-Branch965_plot+ theme(legend.position = "none") #remove legend
Branch966_plot<-Branch966_plot+ theme(legend.position = "none") #remove legend
Branch967_plot<-Branch967_plot+ theme(legend.position = "none") #remove legend
Branch968_plot<-Branch968_plot+ theme(legend.position = "none") #remove legend
Branch969_plot<-Branch969_plot+ theme(legend.position = "none") #remove legend
Branch971_plot<-Branch971_plot+ theme(legend.position = "none") #remove legend
Branch974_plot<-Branch974_plot+ theme(legend.position = "none") #remove legend
Branch976_plot<-Branch976_plot+ theme(legend.position = "none") #remove legend
Branch978_plot<-Branch978_plot+ theme(legend.position = "none") #remove legend
Branch983_plot<-Branch983_plot+ theme(legend.position = "none") #remove legend
 Branch984_plot<-Branch984_plot+ theme(legend.position = "none") #remove legend



grid.arrange(Branch945_plot, Branch957_plot, Branch959_plot,Branch960_plot, 
             Branch961_plot,Branch962_plot,Branch963_plot,Branch964_plot,
             Branch965_plot,Branch966_plot,Branch967_plot,Branch968_plot,
             Branch969_plot,Branch971_plot,Branch974_plot,Branch976_plot,
             Branch978_plot,Branch983_plot,Branch984_plot,nrow = 5)
```

# 2022 Jan 13
Spatial autocorr stats

Within a branch, physical dist vs dist matrix dists, then run mantel to look for correlation.
Try to create a df with pairs of points, only keep 
```{r}
#make a copy of the dataframe
# mData.metadata<-subset(mData.sd, select=c("Distance.to.base","ATTRIBUTE_Sample_Name","ATTRIBUTE_Colony_number","ATTRIBUTE_Branch_Letter")) 

#write.csv(mData.metadata,"mData.metadata2.csv")

#create dist matrix wtih physical and bc dist (did in excel)
Dist.compare.df<-read.csv("distances.csv")

#run pearson correlation 
library(ggpubr)
shapiro.test(Dist.compare.df$bc.dist)
shapiro.test(Dist.compare.df$physical.dist)

ggqqplot(Dist.compare.df$bc.dist, ylab = "bc.dist")
# wt
ggqqplot(Dist.compare.df$physical.dist, ylab = "physical.dist")

cor(Dist.compare.df$bc.dist, Dist.compare.df$physical.dist, method = c("pearson")) # 0.1669614 = corr 


#coeefficient Pearsons
cor.test(Dist.compare.df$bc.dist, Dist.compare.df$physical.dist, method = c("pearson"))

#coeefficient Pearsons rank-based bc no normal
cor.test(Dist.compare.df$bc.dist, Dist.compare.df$physical.dist, method = c("kendall"))


#Transform the data
Dist.compare.df2<-Dist.compare.df #make a copy
hist(Dist.compare.df2$bc.dist)
hist(Dist.compare.df2$physical.dist)

Dist.compare.df2$physical.dist_log<-log10(1+Dist.compare.df2$physical.dist)
hist(Dist.compare.df2$physical.dist_log)
Dist.compare.df2$physical.dist_sq<-sqrt(Dist.compare.df2$physical.dist)
hist(Dist.compare.df2$physical.dist_sq)
Dist.compare.df2$physical.dist_cub<-Dist.compare.df2$physical.dist^(1/3)
hist(Dist.compare.df2$physical.dist_cub)
ggqqplot(Dist.compare.df2$physical.dist_sq, ylab = "physical.dists")


Dist.compare.df$col.branch<-paste(Dist.compare.df$Coral, Dist.compare.df$Branch)

# all together
alldist_plot<-ggplot(data=Dist.compare.df, aes(x=physical.dist, y=bc.dist, color=col.branch)) +
  geom_point(aes(color=col.branch), size=4, show.legend = T, alpha=I(0.1))+
geom_smooth(method = "lm", fill=NA, size=.5)+  #
  #scale_color_manual(values=c("dodgerblue3", "darkred"))+
  xlab("physical.dist") + #Label the X Axis
  #ylim(0, 11) + #set Y limits
  #scale_y_log10()+
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        text = element_text(size=18),  # set element text
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=1))+
        theme(aspect.ratio=1)+
  ylab(expression(paste("bc.dist"))) +
  ggtitle("all")+
  theme(plot.title = element_text(size=20, face = "italic"));alldist_plot

#single line
alldist_plot<-ggplot(data=Dist.compare.df, aes(x=physical.dist, y=bc.dist)) +
  geom_point(aes(color=Coral), size=4, show.legend = T, alpha=I(0.1))+
geom_smooth(method = "lm", fill=NA, size=.5)+  #
  #scale_color_manual(values=c("dodgerblue3", "darkred"))+
  xlab("physical.dist") + #Label the X Axis
  #ylim(0, 11) + #set Y limits
  #scale_y_log10()+
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        text = element_text(size=18),  # set element text
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=1))+
        theme(aspect.ratio=1)+
  ylab(expression(paste("bc.dist"))) +
  ggtitle("all")+
  theme(plot.title = element_text(size=20, face = "italic"));alldist_plot

#r2 and p for the line
lineMod<-lm(bc.dist~physical.dist, data=Dist.compare.df)
anova(lineMod)







#all as boxplot

boxplot(bc.dist~physical.dist, data=Dist.compare.df, notch=TRUE,
  col=(c("gold","darkgreen")),
  main="Spatial distance", xlab="Location")


Dist.compare.df$PolypZones<-as.factor(Dist.compare.df$PolypZones)

Dist.compare.df$PolypZones <- factor(Dist.compare.df$PolypZones, levels = c("Base", "Base-4", "4_5", "Base-5", "5_6","4_6", "Base-6"))

#means
distcompare.means<- ddply(Dist.compare.df, c(  "PolypZones"), summarise, #summarize cell counts
                 N    = length(bc.dist[!is.na(bc.dist)]), #calculate the length of the data frame, excluding NA’s *this part is important*
                 mean = mean(bc.dist, na.rm=TRUE), #calculate mean of response variable, removing NA's
                 sd   = sd(bc.dist, na.rm=TRUE), #calculate standard deviation
                 se   = sd / sqrt(N), #calculate standard error
                 max = max(bc.dist, na.rm=TRUE) #calculate max, could also calculate min () if desired
)
distcompare.means

ggplot(distcompare.means, aes(x=PolypZones, y=mean, fill=PolypZones)) + 
  ylim(0.16,.3)+
  geom_point(aes(fill=PolypZones))+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                width=0.1, show.legend = F)+
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        text = element_text(size=18),  # set element text
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=1))+
        theme(aspect.ratio=1)+
  ylab(expression(paste("bc.dist"))) +
    geom_boxplot(outlier.shape = NA)

quantiles_95 <- function(x) {
  r <- quantile(x, probs=c(0.05, 0.25, 0.5, 0.75, 0.95))
  names(r) <- c("ymin", "lower", "middle", "upper", "ymax")
  r
}

ggplot(Dist.compare.df, aes(x=bc.dist, y=PolypZones, fill=PolypZones)) +
    guides(fill=F) +
    coord_flip() +
    stat_summary(fun.data = quantiles_95, geom="boxplot")



#trying again

Dist.compare.dfTEST <- Dist.compare.df %>% 
    group_by(PolypZones) %>% 
    summarise(
        y0 = quantile(bc.dist, 0.05), 
        y25 = quantile(bc.dist, 0.25), 
        y50 = mean(bc.dist), 
        y75 = quantile(bc.dist, 0.75), 
        y100 = quantile(bc.dist, 0.95))


##### This is the one used in Fig 2 panel D)
# boxplot using the mean for the middle and 95% quantiles for the whiskers
ggplot(Dist.compare.dfTEST, aes(PolypZones, fill = as.factor(PolypZones))) +
    geom_boxplot(
        aes(ymin = y0, lower = y25, middle = y50, upper = y75, ymax = y100),
        stat = "identity"
    ) + 
    labs(y = "len", title = "Boxplot with Mean Middle Line") + 
    theme(plot.title = element_text(hjust = 0.5))









#with st dev
ggplot(distcompare.means, aes(x=PolypZones, y=mean, fill=PolypZones)) + 
 # ylim(0.16,.3)+
  geom_point(aes(fill=PolypZones))+
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd),
                width=0.1, show.legend = F)+
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        text = element_text(size=18),  # set element text
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=1))+
        theme(aspect.ratio=1)+
  ylab(expression(paste("bc.dist"))) +
    geom_boxplot(outlier.shape = NA)







ggplot(Dist.compare.df, aes(x=PolypZones, y=bc.dist, fill=PolypZones)) + 
  #ylim(0.05,.3)+
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        text = element_text(size=18),  # set element text
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=1))+
        theme(aspect.ratio=1)+
  ylab(expression(paste("bc.dist"))) +
    geom_boxplot() # Or outlier.size = -1 instead of outlier.shape = NA)


#try mean and not median
ggplot(Dist.compare.df, aes(x=PolypZones, y=bc.dist, fill=PolypZones)) + 
  #(0,.6)+
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        text = element_text(size=18),  # set element text
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=1))+
        theme(aspect.ratio=1)+
  ylab(expression(paste("bc.dist"))) +
    geom_boxplot(outlier.shape = NA,coef = 0) # Or outlier.size = -1 instead of outlier.shape = NA)

plot <- ggplot(data = Dist.compare.df, aes(y = bc.dist, x = PolypZones)) + 
  geom_boxplot(fatten = NULL,outlier.shape = NA) +
  stat_summary(fun = mean, geom = "errorbar", aes(ymax = ..y.., ymin = ..y..),
               width = 0.75, size = 1, linetype = "solid")
print(plot)

# plot
p <- ggplot(Dist.compare.df, aes(x=PolypZones, y=bc.dist, fill=PolypZones)) +
    geom_boxplot(alpha=0.7) +
    stat_summary(fun=mean, geom="point", shape=20, size=14, color="red", fill="red") +
    theme(legend.position="none") +
    scale_fill_brewer(palette="Set1");p



```
plots by colony
```{r}
#by colony

#945
Dist.compare.df.945<-subset(Dist.compare.df, Colony.id=="945")

p945_plot<-ggplot(data=Dist.compare.df.945, aes(x=physical.dist, y=bc.dist)) +
  geom_point(aes(color=Branch.id, shape=Branch.id), size=4, show.legend = T)+
  #geom_line(aes(color=Branch.id), linetype=2, show.legend = T)+
  #scale_color_manual(values=c("dodgerblue3", "darkred"))+
  xlab("physical.dist") + #Label the X Axis
  #ylim(0, 11) + #set Y limits
  #scale_y_log10()+
      geom_smooth(method='lm')+
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        text = element_text(size=18),  # set element text
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=1))+
        theme(aspect.ratio=1)+
  ylab(expression(paste("bc.dist"))) +
  ggtitle("945")+
  theme(plot.title = element_text(size=20, face = "italic"));p945_plot 

    

#957
Dist.compare.df.957<-subset(Dist.compare.df, Colony.id=="957")

p957_plot<-ggplot(data=Dist.compare.df.957, aes(x=physical.dist, y=bc.dist)) +
  geom_point(aes(color=Branch.id, shape=Branch.id), size=4, show.legend = T)+
  #geom_line(aes(color=Branch.id), linetype=2, show.legend = T)+
  #scale_color_manual(values=c("dodgerblue3", "darkred"))+
      geom_smooth(method='lm')+
  xlab("physical.dist") + #Label the X Axis
  #ylim(0, 11) + #set Y limits
  #scale_y_log10()+
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        text = element_text(size=18),  # set element text
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=1))+
        theme(aspect.ratio=1)+
  ylab(expression(paste("bc.dist"))) +
  ggtitle("957")+
  theme(plot.title = element_text(size=20, face = "italic"));p957_plot 

#959
Dist.compare.df.959<-subset(Dist.compare.df, Colony.id=="959")

p959_plot<-ggplot(data=Dist.compare.df.959, aes(x=physical.dist, y=bc.dist)) +
  geom_point(aes(color=Branch.id, shape=Branch.id), size=4, show.legend = T)+
    geom_smooth(method='lm')+
  #geom_line(aes(color=Branch.id), linetype=2, show.legend = T)+
  #scale_color_manual(values=c("dodgerblue3", "darkred"))+
  xlab("physical.dist") + #Label the X Axis
  #ylim(0, 11) + #set Y limits
  #scale_y_log10()+
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        text = element_text(size=18),  # set element text
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=1))+
        theme(aspect.ratio=1)+
  ylab(expression(paste("bc.dist"))) +
  ggtitle("959")+
  theme(plot.title = element_text(size=20, face = "italic"));p959_plot 

#960
Dist.compare.df.960<-subset(Dist.compare.df, Colony.id=="960")

p960_plot<-ggplot(data=Dist.compare.df.960, aes(x=physical.dist, y=bc.dist)) +
  geom_point(aes(color=Branch.id, shape=Branch.id), size=4, show.legend = T)+
    geom_smooth(method='lm')+
  #geom_line(aes(color=Branch.id), linetype=2, show.legend = T)+
  #scale_color_manual(values=c("dodgerblue3", "darkred"))+
  xlab("physical.dist") + #Label the X Axis
  #ylim(0, 11) + #set Y limits
  #scale_y_log10()+
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        text = element_text(size=18),  # set element text
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=1))+
        theme(aspect.ratio=1)+
  ylab(expression(paste("bc.dist"))) +
  ggtitle("960")+
  theme(plot.title = element_text(size=20, face = "italic"));p960_plot 

#961
Dist.compare.df.961<-subset(Dist.compare.df, Colony.id=="961")

p961_plot<-ggplot(data=Dist.compare.df.961, aes(x=physical.dist, y=bc.dist)) +
  geom_point(aes(color=Branch.id, shape=Branch.id), size=4, show.legend = T)+
    geom_smooth(method='lm')+
  #geom_line(aes(color=Branch.id), linetype=2, show.legend = T)+
  #scale_color_manual(values=c("dodgerblue3", "darkred"))+
  xlab("physical.dist") + #Label the X Axis
  #ylim(0, 11) + #set Y limits
  #scale_y_log10()+
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        text = element_text(size=18),  # set element text
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=1))+
        theme(aspect.ratio=1)+
  ylab(expression(paste("bc.dist"))) +
  ggtitle("961")+
  theme(plot.title = element_text(size=20, face = "italic"));p961_plot 


#962
Dist.compare.df.962<-subset(Dist.compare.df, Colony.id=="962")

p962_plot<-ggplot(data=Dist.compare.df.962, aes(x=physical.dist, y=bc.dist)) +
  geom_point(aes(color=Branch.id, shape=Branch.id), size=4, show.legend = T)+
    geom_smooth(method='lm')+
  #geom_line(aes(color=Branch.id), linetype=2, show.legend = T)+
  #scale_color_manual(values=c("dodgerblue3", "darkred"))+
  xlab("physical.dist") + #Label the X Axis
  #ylim(0, 11) + #set Y limits
  #scale_y_log10()+
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        text = element_text(size=18),  # set element text
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=1))+
        theme(aspect.ratio=1)+
  ylab(expression(paste("bc.dist"))) +
  ggtitle("962")+
  theme(plot.title = element_text(size=20, face = "italic"));p962_plot 



#963
Dist.compare.df.963<-subset(Dist.compare.df, Colony.id=="963")

p963_plot<-ggplot(data=Dist.compare.df.963, aes(x=physical.dist, y=bc.dist)) +
  geom_point(aes(color=Branch.id, shape=Branch.id), size=4, show.legend = T)+
    geom_smooth(method='lm')+
  #geom_line(aes(color=Branch.id), linetype=2, show.legend = T)+
  #scale_color_manual(values=c("dodgerblue3", "darkred"))+
  xlab("physical.dist") + #Label the X Axis
  #ylim(0, 11) + #set Y limits
  #scale_y_log10()+
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        text = element_text(size=18),  # set element text
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=1))+
        theme(aspect.ratio=1)+
  ylab(expression(paste("bc.dist"))) +
  ggtitle("963")+
  theme(plot.title = element_text(size=20, face = "italic"));p963_plot 

#964
Dist.compare.df.964<-subset(Dist.compare.df, Colony.id=="964")

p964_plot<-ggplot(data=Dist.compare.df.964, aes(x=physical.dist, y=bc.dist)) +
  geom_point(aes(color=Branch.id, shape=Branch.id), size=4, show.legend = T)+
    geom_smooth(method='lm')+
  #geom_line(aes(color=Branch.id), linetype=2, show.legend = T)+
  #scale_color_manual(values=c("dodgerblue3", "darkred"))+
  xlab("physical.dist") + #Label the X Axis
  #ylim(0, 11) + #set Y limits
  #scale_y_log10()+
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        text = element_text(size=18),  # set element text
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=1))+
        theme(aspect.ratio=1)+
  ylab(expression(paste("bc.dist"))) +
  ggtitle("964")+
  theme(plot.title = element_text(size=20, face = "italic"));p964_plot 

#965
Dist.compare.df.965<-subset(Dist.compare.df, Colony.id=="965")

p965_plot<-ggplot(data=Dist.compare.df.965, aes(x=physical.dist, y=bc.dist)) +
  geom_point(aes(color=Branch.id, shape=Branch.id), size=4, show.legend = T)+
    geom_smooth(method='lm')+
  #geom_line(aes(color=Branch.id), linetype=2, show.legend = T)+
  #scale_color_manual(values=c("dodgerblue3", "darkred"))+
  xlab("physical.dist") + #Label the X Axis
  #ylim(0, 11) + #set Y limits
  #scale_y_log10()+
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        text = element_text(size=18),  # set element text
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=1))+
        theme(aspect.ratio=1)+
  ylab(expression(paste("bc.dist"))) +
  ggtitle("965")+
  theme(plot.title = element_text(size=20, face = "italic"));p965_plot 

#966
Dist.compare.df.966<-subset(Dist.compare.df, Colony.id=="966")

p966_plot<-ggplot(data=Dist.compare.df.966, aes(x=physical.dist, y=bc.dist)) +
  geom_point(aes(color=Branch.id, shape=Branch.id), size=4, show.legend = T)+
    geom_smooth(method='lm')+
  #geom_line(aes(color=Branch.id), linetype=2, show.legend = T)+
  #scale_color_manual(values=c("dodgerblue3", "darkred"))+
  xlab("physical.dist") + #Label the X Axis
  #ylim(0, 11) + #set Y limits
  #scale_y_log10()+
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        text = element_text(size=18),  # set element text
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=1))+
        theme(aspect.ratio=1)+
  ylab(expression(paste("bc.dist"))) +
  ggtitle("966")+
  theme(plot.title = element_text(size=20, face = "italic"));p966_plot 

#967
Dist.compare.df.967<-subset(Dist.compare.df, Colony.id=="967")

p967_plot<-ggplot(data=Dist.compare.df.967, aes(x=physical.dist, y=bc.dist)) +
  geom_point(aes(color=Branch.id, shape=Branch.id), size=4, show.legend = T)+
    geom_smooth(method='lm')+
  #geom_line(aes(color=Branch.id), linetype=2, show.legend = T)+
  #scale_color_manual(values=c("dodgerblue3", "darkred"))+
  xlab("physical.dist") + #Label the X Axis
  #ylim(0, 11) + #set Y limits
  #scale_y_log10()+
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        text = element_text(size=18),  # set element text
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=1))+
        theme(aspect.ratio=1)+
  ylab(expression(paste("bc.dist"))) +
  ggtitle("967")+
  theme(plot.title = element_text(size=20, face = "italic"));p967_plot 

#968
Dist.compare.df.968<-subset(Dist.compare.df, Colony.id=="968")

p968_plot<-ggplot(data=Dist.compare.df.968, aes(x=physical.dist, y=bc.dist)) +
  geom_point(aes(color=Branch.id, shape=Branch.id), size=4, show.legend = T)+
    geom_smooth(method='lm')+
  #geom_line(aes(color=Branch.id), linetype=2, show.legend = T)+
  #scale_color_manual(values=c("dodgerblue3", "darkred"))+
  xlab("physical.dist") + #Label the X Axis
  #ylim(0, 11) + #set Y limits
  #scale_y_log10()+
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        text = element_text(size=18),  # set element text
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=1))+
        theme(aspect.ratio=1)+
  ylab(expression(paste("bc.dist"))) +
  ggtitle("968")+
  theme(plot.title = element_text(size=20, face = "italic"));p968_plot 

#969
Dist.compare.df.969<-subset(Dist.compare.df, Colony.id=="969")

p969_plot<-ggplot(data=Dist.compare.df.969, aes(x=physical.dist, y=bc.dist)) +
  geom_point(aes(color=Branch.id, shape=Branch.id), size=4, show.legend = T)+
    geom_smooth(method='lm')+
  #geom_line(aes(color=Branch.id), linetype=2, show.legend = T)+
  #scale_color_manual(values=c("dodgerblue3", "darkred"))+
  xlab("physical.dist") + #Label the X Axis
  #ylim(0, 11) + #set Y limits
  #scale_y_log10()+
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        text = element_text(size=18),  # set element text
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=1))+
        theme(aspect.ratio=1)+
  ylab(expression(paste("bc.dist"))) +
  ggtitle("969")+
  theme(plot.title = element_text(size=20, face = "italic"));p969_plot 

#971
Dist.compare.df.971<-subset(Dist.compare.df, Colony.id=="971")

p971_plot<-ggplot(data=Dist.compare.df.971, aes(x=physical.dist, y=bc.dist)) +
  geom_point(aes(color=Branch.id, shape=Branch.id), size=4, show.legend = T)+
    geom_smooth(method='lm')+
  #geom_line(aes(color=Branch.id), linetype=2, show.legend = T)+
  #scale_color_manual(values=c("dodgerblue3", "darkred"))+
  xlab("physical.dist") + #Label the X Axis
  #ylim(0, 11) + #set Y limits
  #scale_y_log10()+
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        text = element_text(size=18),  # set element text
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=1))+
        theme(aspect.ratio=1)+
  ylab(expression(paste("bc.dist"))) +
  ggtitle("971")+
  theme(plot.title = element_text(size=20, face = "italic"));p971_plot 

#974
Dist.compare.df.974<-subset(Dist.compare.df, Colony.id=="974")

p974_plot<-ggplot(data=Dist.compare.df.974, aes(x=physical.dist, y=bc.dist)) +
  geom_point(aes(color=Branch.id, shape=Branch.id), size=4, show.legend = T)+
    geom_smooth(method='lm')+
  #geom_line(aes(color=Branch.id), linetype=2, show.legend = T)+
  #scale_color_manual(values=c("dodgerblue3", "darkred"))+
  xlab("physical.dist") + #Label the X Axis
  #ylim(0, 11) + #set Y limits
  #scale_y_log10()+
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        text = element_text(size=18),  # set element text
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=1))+
        theme(aspect.ratio=1)+
  ylab(expression(paste("bc.dist"))) +
  ggtitle("974")+
  theme(plot.title = element_text(size=20, face = "italic"));p974_plot 

#976
Dist.compare.df.976<-subset(Dist.compare.df, Colony.id=="976")

p976_plot<-ggplot(data=Dist.compare.df.976, aes(x=physical.dist, y=bc.dist)) +
  geom_point(aes(color=Branch.id, shape=Branch.id), size=4, show.legend = T)+
    geom_smooth(method='lm')+
  #geom_line(aes(color=Branch.id), linetype=2, show.legend = T)+
  #scale_color_manual(values=c("dodgerblue3", "darkred"))+
  xlab("physical.dist") + #Label the X Axis
  #ylim(0, 11) + #set Y limits
  #scale_y_log10()+
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        text = element_text(size=18),  # set element text
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=1))+
        theme(aspect.ratio=1)+
  ylab(expression(paste("bc.dist"))) +
  ggtitle("976")+
  theme(plot.title = element_text(size=20, face = "italic"));p976_plot 

#978
Dist.compare.df.978<-subset(Dist.compare.df, Colony.id=="978")

p978_plot<-ggplot(data=Dist.compare.df.978, aes(x=physical.dist, y=bc.dist)) +
  geom_point(aes(color=Branch.id, shape=Branch.id), size=4, show.legend = T)+
    geom_smooth(method='lm')+
  #geom_line(aes(color=Branch.id), linetype=2, show.legend = T)+
  #scale_color_manual(values=c("dodgerblue3", "darkred"))+
  xlab("physical.dist") + #Label the X Axis
  #ylim(0, 11) + #set Y limits
  #scale_y_log10()+
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        text = element_text(size=18),  # set element text
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=1))+
        theme(aspect.ratio=1)+
  ylab(expression(paste("bc.dist"))) +
  ggtitle("978")+
  theme(plot.title = element_text(size=20, face = "italic"));p978_plot 

#983
Dist.compare.df.983<-subset(Dist.compare.df, Colony.id=="983")

p983_plot<-ggplot(data=Dist.compare.df.983, aes(x=physical.dist, y=bc.dist)) +
  geom_point(aes(color=Branch.id, shape=Branch.id), size=4, show.legend = T)+
    geom_smooth(method='lm')+
  #geom_line(aes(color=Branch.id), linetype=2, show.legend = T)+
  #scale_color_manual(values=c("dodgerblue3", "darkred"))+
  xlab("physical.dist") + #Label the X Axis
  #ylim(0, 11) + #set Y limits
  #scale_y_log10()+
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        text = element_text(size=18),  # set element text
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=1))+
        theme(aspect.ratio=1)+
  ylab(expression(paste("bc.dist"))) +
  ggtitle("983")+
  theme(plot.title = element_text(size=20, face = "italic"));p983_plot 

#984
Dist.compare.df.984<-subset(Dist.compare.df, Colony.id=="984")

p984_plot<-ggplot(data=Dist.compare.df.984, aes(x=physical.dist, y=bc.dist)) +
  geom_point(aes(color=Branch.id, shape=Branch.id), size=4, show.legend = T)+
    geom_smooth(method='lm')+
  #geom_line(aes(color=Branch.id), linetype=2, show.legend = T)+
  #scale_color_manual(values=c("dodgerblue3", "darkred"))+
  xlab("physical.dist") + #Label the X Axis
  #ylim(0, 11) + #set Y limits
  #scale_y_log10()+
  theme_bw() + #Set the background color
  theme(axis.line = element_line(color = 'black'), #Set the axes color
        axis.title=element_text(size=14,face="bold"), #Set axis format
        panel.border = element_blank(), #Set the border
        panel.grid.major = element_blank(), #Set the major gridlines
        panel.grid.minor = element_blank(), #Set the minor gridlines
        text = element_text(size=18),  # set element text
        plot.background =element_blank(), #Set the plot background
        legend.key = element_blank()) + #Set plot legend key
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(colour = "black", size=1))+
        theme(aspect.ratio=1)+
  ylab(expression(paste("bc.dist"))) +
  ggtitle("984")+
  theme(plot.title = element_text(size=20, face = "italic"));p984_plot 

#plot all 
# The next 4 plots created will be plotted next to each other - this prins terribly

#grid.arrange(p945_plot, p957_plot, p959_plot,p960_plot,p961_plot,
          # p962_plot,p963_plot,p964_plot,
          # p965_plot,p966_plot,p967_plot,p968_plot,
          #    p969_plot,p971_plot,
          #    p974_plot,p976_plot,p978_plot,p983_plot,p984_plot, nrow = 5)
```
 Models on distance compare
 physical.dist_sq for models
 
```{r}
Dist.compare.df$Coral<-as.factor(as.character(Dist.compare.df$Coral))
mod<-lmer(bc.dist~physical.dist + (1|Coral) + (1|Coral:Branch), data=Dist.compare.df )
anova(mod)
  shapiro.test(Dist.compare.df$bc.dist) #not normal
  shapiro.test(Dist.compare.df$physical.dist) #not normal
ggqqplot(Dist.compare.df$bc.dist, ylab = "bc.dist")
# wt
ggqqplot(Dist.compare.df$physical.dist, ylab = "physical.dist")


#Transform the data
Dist.compare.df<-Dist.compare.df #make a copy
hist(Dist.compare.df$bc.dist)
  Dist.compare.df$bc.dist_log<-log10(1+Dist.compare.df$bc.dist)
  hist(Dist.compare.df$bc.dist_log)
  Dist.compare.df$bc.dist_sqrt<-sqrt(Dist.compare.df$bc.dist)
  hist(Dist.compare.df$bc.dist_sqrt) #looks good use this


hist(Dist.compare.df$physical.dist)
  Dist.compare.df$physical.dist_log<-log10(1+Dist.compare.df$physical.dist)
  hist(Dist.compare.df$physical.dist_log)
ggqqplot(Dist.compare.df$physical.dist_log, ylab = "phys.dist")

Dist.compare.df$physical.dist_t<-(Dist.compare.df$physical.dist)^1/3
hist(Dist.compare.df$physical.dist_t)
ggqqplot(Dist.compare.df$physical.dist_t, ylab = "phys.dist")


Dist.compare.df$physical.dist_sq<-sqrt(Dist.compare.df$physical.dist)
hist(Dist.compare.df$physical.dist_sq)
Dist.compare.df$physical.dist_cub<-Dist.compare.df$physical.dist^(1/3)
hist(Dist.compare.df$physical.dist_cub)
ggqqplot(Dist.compare.df$physical.dist_sq, ylab = "physical.dists")


Dist.mod<-lmer(bc.dist_sqrt~physical.dist_t + (1|Coral) + (1|Coral:Branch), data=Dist.compare.df )
anova(Dist.mod)  # differ
qqPlot(residuals(Dist.mod))  # residuals look good
#physical.dist 0.23483 0.23483     1 817.58  32.727 1.488e-08 ***

# mod<-lm(bc.dist~physical.dist, data=Dist.compare.df)#of not transformed data
# anova(mod)
# qqPlot(residuals(mod))  

```

```{r}
### By polyp zones FIG1H! #########
mod<- lmer(bc.dist_sqrt~PolypZones + (1|Coral/Branch) , data=Dist.compare.df )
anova(mod)
summary(mod)
r.squaredGLMM(mod)
qqPlot(residuals(mod)) # looks good
emm<-emmeans(mod,  ~ PolypZones, adjust="Tukey")
cld(emm)
pairs(emm)

#use this non-parametric Kruskal-Wallis test
kruskal.test(bc.dist ~ PolypZones, data = Dist.compare.df)
pairwise.wilcox.test(Dist.compare.df$bc.dist_sqrt, Dist.compare.df$PolypZones,
                 p.adjust.method = "BH")


#mod<- lmer(bc.dist~PolypZones + (1|Coral) + (1|Branch:Coral) , data=Dist.compare.df )
#anova(mod)
#emm<-emmeans(mod,  ~ PolypZones, adjust="Tukey")
#cld(emm)
#pairs(emm)




```

